Road traffic sound level estimation from realistic urban sound mixtures by Non-negative Matrix Factorization

### 2018-04-06 paper/articleNmfTrafficSimScene2018.tex
jfp corrections

In this paper, a Non-negative Matrix Factorization (NMF) framework is developed to estimate road traffic noise levels within urban sound scenes. To evaluate the performance of the proposed approach, a synthetic corpus of sound scenes is designed to cover most common soundscape settings. The realism of the produced scenes is validated through a perceptual test, forming a protocol that both reproduces the sensor network outputs, and in which the actual occurrence and sound level of each source is known.

Several variants of NMF are tested. The threshold initialized NMF appears to be the most reliable approach, allowing road traffic noise level estimation with average errors of less than 1.3 dB over the tested corpus of sound scenes.

@@

\begin{keyword}
non-negative matrix factorization \sep urban sound environment \sep road traffic sound level estimation
\end{keyword}

@@

In response to the growing demand from urban dwellers for a better environment, noise mapping has been recommended as a tool to tackle noise pollution. The enactment of the European Directive 2002/EC/49 makes such maps mandatory to cities over 100 000 inhabitants. Those maps play an important informative role, establishing the distribution of the sound levels all over the cities as well as the estimation of the number of city dwellers exposed to high sound level ($>$ 55 dB(A)) \cite{nugent2014noise}. Road traffic concentrates particular attention as it is the main urban source of noise annoyance. Road traffic noise maps are typically built from data collection that consist of traffic data collected on the main roads (flow rates, mean speeds and heavy vehicle ratio) and urban geographic data (building heights and location, topology, ground surfaces, etc.). Follows sound emission and sound propagation computational techniques, resulting in the production of the two indicators equivalent A-weighted sound levels, $L_{DEN}$ (\textit{Day-Evening-Night}) and $L_N$ (\textit{Night}) \cite{kephalopoulos2012common}. This procedure also enables drawing up action plans to reduce the noise exposure. Despite their unanimously recognized interest, noise maps suffer some limitations. The computing cost required to produce noise maps at the city scale calls simplifications of the numerical tools and the simulation models that both generate uncertainties \cite{van_leeuwen_noise_2015}. Data collection is itself also a vector of uncertainty. Moreover, the produced aggregated indicators do not model the sound levels evolution due to the traffic variations throughout the day.

@@

All these measurement protocols allows the combination of measures and predictions to improve the accuracy of the produced noise maps. Traffic noise maps and measurements were compared on restrictive areas in \cite{lefebvre2017traffic} and \cite{mioduszewski2011noise}. Wei et al. \cite{wei_dynamic_2016} modify the acoustical parameters of the simulation thanks to noise measurements, while  Mallet et al. \cite{ventura2017estimation} call for data assimilation techniques between models and measurements to reduce the uncertainty of the produced noise maps. However, these works make the implicit assumption that the noise measurements consist mainly of road traffic. In the aim to improve road traffic noise maps, the use of measurements has first to deal with the challenge to estimate correctly the road traffic sound level.
Even if road traffic is predominant on many urban areas, urban sound environments are composed of many different overlapping sound sources (passing cars, voices, footsteps, car horn, whistling birds \dots), what makes the task of estimating correctly the traffic sound level within an urban sound mixture not trivial.

Many works have dealt with the detection \cite{luitel2016sound} or the recognition \cite{defreville_automatic_2006} of sound events in environmental sound scenes. In these cases, a two-step scheme is followed where audio samples are described with a set of features (Mel Frequency Cepstral Coefficient, MPEG-7 descriptors \dots) and classified them with the help of a classifiers (Gaussian Mixtures Models, Artificial Neural Network \dots) \cite{chu2008environmental, cowling_comparison_2003}. The classifiers is learnt from a learning database and are next applied on a test database to validate the algorithms.
Dedicated to the traffic, in \cite{socoro_anomalous_2017}, an Anomalous Event Detection, based on MFCC features, is proposed with the specific aim to improve the traffic sound estimation. It is based on the detection of unwanted sound events in order to discard them.

An other approach, followed in this paper, is to consider the blind source separation paradigm which consists in the extraction of a specific signal inside a set of mixed signals, see Figure \ref{fig:source_separation}. From the different existing methods, Non-negative Matrix Factorization (NMF) \cite{lee_learning_1999}, appears to be a relevant method for monophonic sensor networks. Many applications can be found for musical \cite{smaragdis_non-negative_2003, benetos2006musical} and speech \cite{wilson_speech_2008, mysore2011non} contents. Dedicated to sound separation with environmental sound, Immani and Kasa\"i \cite{satoshi_innami_nmf-based_2012} used NMF in a two steps sound separation with the help of time variant gain features. Dedicated to the traffic sound separation, a first study \cite{gloaguen2018Estimation} has been conducted, in which diverse NMF estimation rules are compared, namely the supervised, the semi-supervised, and the threshold initialized NMF, have been applied on a large set of simulated sound scenes. This corpus mixes 6 sound categories (\textit{alert, animals, climate, humans, mechanics, transportation}) with a traffic component calibrated to different sound levels, according to the other sound classes (in the rest of the document, these sound classes are resumed as the \textit{interfering} sound class), to obtain variable traffic predominance. 
The diversity of this corpus was made to assess the performances and the limits of each NMF formula. 
However, if this study reveals the interest of NMF for urban sound environments, the assessment of its performance on a corpus of realistic sound scenes must be carried out in order to implement it on a sensor network. 
Design urban sound mixtures makes it possible to access to many acoustic proprieties as the onset and offset time and the sound level of each sound class and especially the traffic component. The realistic aspect of such a corpus is essential to obtain sound scenes similar to recordings and then validate NMF performances. However, like all simulated process, the issue of its realism has to be resolved.

In this paper, a realistic urban sound corpus based on annotated urban recordings, and whose degree of realism is validated through a perceptual test, is designed in order to estimate the traffic sound level with the help of the NMF framework.
The different NMF approaches are described in section \ref{part:nmf}. Next, the corpus of urban sound scenes is presented in section \ref{part:urban_scene}, from the sound database built-up to its validation. The experimental protocol and the results are then presented and discussed in section \ref{part:expProtocol} and \ref{part:results}.

@@

The choice of the dimensions is often made such as $F\times K + K \times N < F \times N$ so that NMF can be a low rank approximation. This condition however is not mandatory. When  applying NMF to audio data, $\mathbf{V}$ is usually considered as the magnitude spectrogram obtained by a Short-Time Frequency Transform, $\mathbf{W}$ includes audio spectra and $\mathbf{H}$ is equivalent to the temporal activation of each spectrum, see Figure \ref{fig:exampleNMF}. Because of the non-negativity constraint, only additive combinations between the elements of $\mathbf{W}$ are considered.

\begin{figure}[t]
\includegraphics[width=.9\linewidth]{./figures/schema_introduction_nmf.pdf}
\caption{NMF decomposition of an audio spectrogram composed of 3 elements ($K$ = 3): passing car (a), car horn (b) and whistling bird (c).}
\label{fig:exampleNMF}
\end{figure}

@@

\subsection{Supervised NMF}
The most easiest case of NMF is the one where the sound sources can be known \textit{a priori} and $\mathbf{W}$ can be built directly from audio samples. It leads to \textit{supervised} NMF (SUP-NMF). $\mathbf{H}$ is then the only matrix to estimate and is updated at every iteration (eq. \ref{eq:updateH}) \cite{fevotte_algorithms_2011}.

@@

with $\gamma(\beta) = \frac{1}{2-\beta},$ for $\beta < 1$, $ \gamma(\beta) = 1$, for $\beta \in \left[1,2\right]$ and $\gamma(\beta) = \frac{1}{\beta-1}$ for $\beta > 2$. The product $A\otimes B$ and $A/B$ symbolized the Hadamard product and ratio.

@@

The main issue with the supervised approach is the representational limit imposed by a fixed $\mathbf{W}$. To be completely successful, all the acoustical sources must be considered in it which is not possible in an urban environment. To overcome this issue, semi-supervised NMF (SEM-NMF) \cite{lee_semi-supervised_2010} has been proposed. It consists in decomposing, $\mathbf{W}_{F \times (K+J)}$ into two distinctive matrices: $\mathbf{W} = \left[ \mathbf{W_s}~\mathbf{W_r} \right]$ where $\mathbf{W_s}_{F \times K}$ is a fixed part of $\mathbf{W}$ composed of audio spectra and $ \mathbf{W_r}_{F \times J}$, a mobile part which is updated, see eq. \ref{eq:W_r_SS}. Thus it is possible to include in it elements not present in $\mathbf{W_s}$. The dimension of $\mathbf{W_r}$ is set up as $J << K$ in order to best considered the sound source present in $\mathbf{W_s}$. $\mathbf{H}$ is then also decomposed in two matrices, $\mathbf{H}_{(K+J) \times N} = \genfrac[]{0pt}{0}{\mathbf{H_s}}{\mathbf{H_r}}$. The eq. \ref{eq:nmf} becomes

@@

$\mathbf{H_r}$ and $\mathbf{H_s}$ are updated separately, see eq. \ref{eq:H_r_SS} \ref{eq:H_s_SS}.

@@

In this study, $\mathbf{W_s}$ is composed of traffic audio spectra, as it is the sound source of interest, to include in $\mathbf{W_r}$ the interfering class that can be present in the urban sound scenes. The traffic signal estimation is next defined by the fixed part,

@@

The addition of the mobile part $\mathbf{W_r}$ gives more flexibility. The representational capability is increased, thus the approach is more adaptive to the different urban sound environments. Applications of SEM-NMF can be found for musical \cite{weninger2012supervised, kitamura_music_2014} and speech content \cite{joder2012real, mysore2011non}.

@@

With this operation, $\mathbf{W_0}$ is oriented to the focused sound source (the road traffic) but also can be adapted to the content of the scene thanks to the updates. After $N$ iterations, each element $k$ of the final dictionary, $\mathbf{W'}$, is compared with its initial value in $\mathbf{W_0}$, in order to identify which element is stayed closed to the traffic component. A cosine similarity $D_{\theta}\left(\mathbf{W_0} \Vert \mathbf{W'} \right)$ is computed for each element $k$ as it is an invariant scale and a bounded method,

@@

where $\mathbf{w}$ is a $k$ element of $\mathbf{W}$ of $F \times 1$ dimensions. When $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$=1, both element are identicals. If $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$=0, the element $\mathbf{w'}$ is fully different from $\mathbf{w_0}$. The extraction of traffic elements in $\mathbf{W'}$ is carried out by a hard thresholding method \cite{donoho1994threshold}. It consists in weighting in a binary way $\mathbf{W'}$ according to $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$ and a threshold value such as:

@@

To summarize, to approximate the audio spectrogram, $\mathbf{V}$ and estimate the traffic component, 3 methods are used which deal differently the \textit{a priori} knowledge on the traffic. SUP-NMF is only based on fixed \textit{traffic} elements and is constrained to used it to obtain $\mathbf{\tilde{V}}_{traffic}$. SEM-NMF completes this dictionary by the add of a mobile dictionary to consider the interfering sound events and then offer more flexibility. Finally, TI-NMF considers first a dictionary composed of \textit{traffic} spectra, as SUP-NMF, but to then update them. This method therefore makes it the most suitable for solving the generalization issues as it is the entire dictionary that can be fully adapted to the sound scenes. The interest of an initiated dictionary is then to focus on the source of interest during the updates.

These above described methods are applied on an evaluation corpus,  composed of simulated sound scenes, in order to compare the estimated sound levels with a non ambiguous reference. This new and more realistic sound corpus is designed by considering realistic urban sound environments with many kind of sound sources mixed.

@@

\begin{figure*}[t]
\centering
\includegraphics[width=0.7\linewidth]{figures/realistic_urban_sound_scene_design.pdf}
\caption{Bloc diagram of the design of evaluation corpus}
\label{fig:transcription}
\end{figure*}

The evaluation corpus is both required to be realistic and non ambiguous in terms of traffic sound level. The former calls for the use of recordings of urban sound environments, and the latter imposes the use of controlled sound sources with known properties: onset and offset time location, and sound level. Indeed, the precise annotation of the sound level of a given source of interest in a polyphonic mixture is not feasible. Two corpora are proposed to design such a corpus of  evaluation. The first, called the reference corpus, is a set of recordings of urban sound environments, from which a sort of transcription is made, \textit{i.e.} which sound source is present from this time to this time, at which sound level. The second, called the elemental corpus, is a set of monophonic recordings of isolated sound sources that are occurs in the scenes of the reference. The elemental corpus is built with still existing sound and with specific recordings, see part \ref{part:simScene}. The evaluation corpus is then designed by replicating the reference corpus according to the transcription with the help of the elemental corpus, see Figure \ref{fig:transcription}.
More precisely, the urban sound scenes are taken from 76 recordings from 2 to 5 min, achieved in the \nth{13} district of Paris (France) at 19 different locations \footnote{Recordings were made as part of the Grafic project funded by Ademe}, which cover four various sound environments (Figure \ref{fig:map_grafic}), see Figure \ref{fig:map_grafic}. A complete description of the experimental protocol can be found in \cite{aumond_modelling_2017}. Two of the 76 recordings are rejected for the analysis because the audio files were corrupted, resulting in 74 valid audio files assumed as representative of the variety of sound environments. The recordings are listened and categorized within four different sound environments, as proposed in \cite{can_describing_2015}: \textit{park} (P, 8 audio files with a cumulative duration of 16min01), \textit{quiet street} (Q, 35 audio files with a cumulative duration of 77min27), \textit{noisy street} (N, 23 audio files with a cumulative duration of 56min10) and \textit{very noisy street} (vN, 8 audio files with a cumulative duration of 21min42). Then, each audio file is transcripted, noticing the start, end time and level of each sound event along with its sound class. This annotation phase makes it possible to obtain simulated sound scenes with the same positions of sound events than the recordings and therefore as close as possible to the realistic scenes.

@@

\subsection{Generation of the evaluation corpus}\label{part:simScene}
The sound scenes are generated with the \textit{SimScene} software\footnote{Open-source project available at: \url{https://bitbucket.org/mlagrange/simscene}}, \cite{rossignol_simscene:_2015}, which is a simulation software generating monaural sound mixtures in wav format with a 44.1 kHz sampling rate from an isolated sound database.
This software have already been used in a wide range of experiments for sound detection algorithm assessment \cite{lafay_new_2014, benetos2016detection}. The \textit{SimScene} software allows for several type of sequencing, from abstract ones (time indexes and amplitudes are drawn from random distributions) to  precise ones, where the time indexes and amplitudes for each event is set by the user. The latter type is considered in this study. As output, \textit{SimScene} generates an audio of the global sound mixture and an audio for each sound class present in the scene, which makes it possible to know their exact contributions in the scene, see Figure \ref{fig:example_simScene}.

\begin{figure}[t]
    \centering
       \includegraphics[width=.8\linewidth]{./figures/exampleSimScene.pdf}
    \caption{Spectrogram of a simple scene created with the \textit{SimScene} software with a sound background (road traffic in green) and 3 sound events (car horn in purple, passing car in red and whistling bird in blue).}
    \label{fig:example_simScene}
\end{figure}

To transcribe the recordings in simulated scenes, a high quality sound database (wav format, 44.1 kHz sampling rate, high \textit{Signal Noise Ratio}) has been built-up from audio samples found online (\textit{freesound.org}) or with the help of an already existing sound database \cite{salamon2014dataset}. The sound database is composed of two categories of sound: the \textit{event} category, which includes 245 brief sound samples considered as salient, with a 1 to 20 seconds duration and classified among 21 sound classes (\textit{ringing bell, whistling bird, car horn, passing car, hammer, barking dog, siren, footstep, metallic noise, voice \dots}) and the \textit{background} (or \textit{texture}) category gathering 154 long duration sounds ($\approx$ 1m30), whose acoustic properties do not vary in time. This category includes among others \textit{whistling bird, crowd noise, rain, children playing in schoolyard, constant traffic noise} \dots~sound classes. Each sound class is composed of multiple samples (\textit{carHorn01.wav, carHorn02.wav \dots}) that are randomly chosen by the \textit{simScene} software to bring diversity.
As the road traffic is the main component in urban environment and is the sound source of interest, recordings of car passages has been made on the Ifsttar's runway. The recordings has been made for 4 cars (Renault Scenic, Renault M\'egane, Renault Clio and Dacia Sandero),  at different speeds and gear ratios. Overall, 103 car passages have been recorded. In order to avoid overfitting issues, the audio samples of the first two cars (Renault Scenic and Renault M\'egane) are included in the \textit{SimScene}'s sound database (50 audio files in total). The last 53 audio samples are dedicated to the dictionary design as part of NMF, see section \ref{part:dictionary_building}. A full description of the recordings can be found in \cite{gloaguen_creation_2017}.

% \ml{ajouter une description sur les problématiques d'overfitting et de generalisation. Cela doit etre un rappel d'un texte ecrit dans la partie nmf. A ce titre, cela pourrait etre une extension de l'etude, on en reparle}.

\begin{table}[t]
\centering
\caption{Mean \textit{Traffic Interfering Ratio} for each sound environment.}
\begin{tabular}{lc}
 & $mTIR$ (dB)\\ \hline
 \textit{park} & -9.10 ($\pm$ 7.35) \\
 \textit{quiet street} & 0.88 ($\pm$ 5.92) \\
 \textit{noisy street} & 6.96 ($\pm$ 5.16) \\
 \textit{very noisy street} & 15.75 ($\pm$ 9.78) \\ \hline
\end{tabular}
\label{tab:mTIR}
\end{table}

With this built-up database, the \textit{SimScene} software and the transcriptions of the recordings, 74 simulated sound scenes are generated, which have the temporal structure of the reference recordings. The sound level of each sound class is adjusted manually on each sound scene to be faithful compared to the recorded scenes. To validate this adjustment, the mean \textit{Traffic Interfering Ratio} ($mTIR$) is calculated, see Table \ref{tab:mTIR}. It expresses, on all the scenes of an sound environment, the mean difference between the equivalent traffic sound levels of each scene, $L_{p,traffic}$, with the sound level of the \textit{interfering} sound class, $L_{p, interfering}$, which gathered all the other sound sources not related to the traffic, see Eq \ref{eq:mTIR}. It quantifies the predominance of the traffic component for the 4 types of sound environments.

>>>>

In this paper, the Non-negative Matrix Factorization (NMF) framework is developed to estimate road traffic noise levels within urban sound scenes. To evaluate the performances of the proposed approach, a synthetic corpus of sound scenes is designed, to cover most common soundscape settings, and whom realism is validated through a perceptual test. The simulated scenes reproduce then the sensor network outputs, in which the actual occurrence and sound level of each source are known.

Several variants of NMF are tested. The proposed approach, named threshold initialized NMF, appears to be the most reliable approach, allowing road traffic noise level estimation with average errors of less than 1.3 dB over the tested corpus of sound scenes.

@@

\begin{keyword}
Non-negative Matrix Factorization \sep urban sound environment \sep road traffic sound level estimation
\end{keyword}

@@

In response to the growing demand from urban dwellers for a better environment, noise mapping has been recommended as a tool to tackle noise pollution. The enactment of the European Directive 2002/EC/49 makes such maps mandatory to cities over 100 000 inhabitants. Those maps play an important informative role, establishing the distribution of the sound levels all over the cities as well as the estimation of the number of city dwellers exposed to high sound level ($>$ 55 dB(A)) \cite{nugent2014noise}. Road traffic concentrates particular attention as it is the main urban source of noise annoyance. Road traffic noise maps are typically built from data collection that consist of traffic data collected on the main roads (flow rates, mean speeds and heavy vehicle ratio) and urban geographic data (building heights and location, topology, ground surfaces \dots). Follows sound emission and sound propagation computational techniques, resulting in the production of the two indicators equivalent A-weighted sound levels, $L_{DEN}$ (\textit{Day-Evening-Night}) and $L_N$ (\textit{Night}) \cite{kephalopoulos2012common}. This procedure also enables drawing up action plans to reduce the noise exposure. Despite their unanimously recognized interest, noise maps suffer from some limitations. The computing cost required to produce noise maps at the city scale calls simplifications of the numerical tools and the simulation models that both generate uncertainties \cite{van_leeuwen_noise_2015}. Data collection is itself also a vector of uncertainty. Moreover, the produced aggregated indicators do not model the sound levels evolution due to the traffic variations throughout the day.

@@

All these measurement protocols allow the combination of measures and predictions to improve the accuracy of the produced noise maps. Traffic noise maps and measurements were compared on restrictive areas in \cite{lefebvre2017traffic} and \cite{mioduszewski2011noise}. Wei et al. \cite{wei_dynamic_2016} modify the acoustical parameters of the simulation thanks to noise measurements, while  Mallet et al. \cite{ventura2017estimation} call for data assimilation techniques between models and measurements to reduce the uncertainty of the produced noise maps. However, these works make the implicit assumption that the noise measurements consist mainly of road traffic. In the aim to improve road traffic noise maps, the use of measurements has first to deal with the challenge to estimate correctly the road traffic sound level.
Even if road traffic is predominant on many urban areas, urban sound environments are composed of many different overlapping sound sources (passing cars, voices, footsteps, car horn, whistling birds \dots), what makes the task of estimating correctly the traffic sound level within an urban sound mixture not trivial.

Many works have dealt with the detection \cite{luitel2016sound} or the recognition \cite{defreville_automatic_2006} of sound events in environmental sound scenes. In these cases, a two-step scheme is followed where audio samples are described with a set of features (Mel Frequency Cepstral Coefficient, MPEG-7 descriptors \dots) and classified with the help of a classifier (Gaussian Mixtures Models, Artificial Neural Network \dots) \cite{chu2008environmental, cowling_comparison_2003}. The classifier is learnt from a learning database and is next applied on a test database to validate the algorithms.
Dedicated to the traffic, in \cite{socoro_anomalous_2017}, an Anomalous Event Detection, based on MFCC features, is proposed with the specific aim to improve the traffic sound estimation. It is based on the detection of unwanted sound events in order to discard them.

An other approach, followed in this paper, is to consider the blind source separation paradigm which consists in the extraction of a specific signal inside a set of mixed signals, see Figure \ref{fig:source_separation}. From the different existing methods, Non-negative Matrix Factorization (NMF) \cite{lee_learning_1999}, appears to be a relevant method for monophonic sensor networks. Many applications can be found for musical \cite{smaragdis_non-negative_2003, benetos2006musical} and speech \cite{wilson_speech_2008, mysore2011non} contents. Dedicated to sound separation with environmental sounds, Immani and Kasa\"i \cite{satoshi_innami_nmf-based_2012} used NMF in a two steps sound separation with the help of time variant gain features. Dedicated to the traffic sound separation, a first study \cite{gloaguen2018Estimation} has been conducted, in which diverse NMF estimation rules are compared, namely the supervised, the semi-supervised, and the threshold initialized NMF, have been applied on a large set of simulated sound scenes. This corpus mixes 6 sound categories (\textit{alert, animals, climate, humans, mechanics, transportation}) with a traffic component calibrated to different sound levels, according to the other sound classes (in the rest of the document, these sound classes, not related to the traffic component, are resumed as the \textit{interfering} sound class), to obtain variable traffic predominance. 
The diversity of this corpus was made to assess the performances and the limits of each NMF formula. 
However, if this study reveals the interest of NMF for urban sound environments, the assessment of its performance on a corpus of realistic sound scenes must be carried out in order to implement it on a sensor network. 
Design urban sound mixtures makes it possible to access to many acoustic properties as the onset and offset time and the sound level of each sound class and especially the traffic component. The realistic aspect of such a corpus is essential to obtain sound scenes similar to recordings and to validate NMF performances. However, like all simulated process, the realism of the scenes must be perceptually verified.

In this paper, an urban sound corpus based on annotated urban recordings, and whose degree of realism is assessed through a perceptual test, is designed in order to estimate the traffic sound level with the help of the NMF framework.
The different NMF approaches are described in section \ref{part:nmf}. Next, the corpus of urban sound scenes is presented in section \ref{part:urban_scene}, from the sound database built-up to its validation. The experimental protocol and the results are then presented and discussed in section \ref{part:expProtocol} and \ref{part:results}.

@@

The choice of the dimensions is often made such as $F\times K + K \times N < F \times N$ so that NMF can be a low rank approximation. This condition however is not mandatory. When  applying NMF to audio data, $\mathbf{V}$ is usually considered as the magnitude spectrogram obtained by a Short-Time Fourier Transform, $\mathbf{W}$ includes audio spectra and $\mathbf{H}$ is equivalent to the temporal activation of each spectrum, see Figure \ref{fig:exampleNMF}. Because of the non-negativity constraint, only additive combinations between the elements of $\mathbf{W}$ are considered.

\begin{figure}[t]
\includegraphics[width=.9\linewidth]{./figures/schema_introduction_nmf.pdf}
\caption{NMF decomposition of an audio spectrogram $V$ composed of 3 elements ($K$ = 3): passing car (a), car horn (b) and whistling bird (c).}
\label{fig:exampleNMF}
\end{figure}

@@

\subsection{Supervised NMF}
The most easiest case of NMF is the one where the sound sources can be known \textit{a priori} and $\mathbf{W}$ can be built directly from audio samples. It leads to \textit{supervised} NMF (SUP-NMF). $\mathbf{H}$ is then the only matrix to estimate and is updated at every iteration (eq. \ref{eq:updateH}) \cite{fevotte_algorithms_2011}:

@@

with $\gamma(\beta) = \frac{1}{2-\beta},$ for $\beta < 1$, $ \gamma(\beta) = 1$, for $\beta \in \left[1,2\right]$ and $\gamma(\beta) = \frac{1}{\beta-1}$ for $\beta > 2$. The $A\otimes B$ and $A/B$ operators represent the Hadamard product and ratio.

@@

The main issue with the supervised approach is the representational limit imposed by a fixed $\mathbf{W}$. To be completely successful, all the acoustical sources must be considered in the basis $\mathbf{W}$ which is not always possible in a complex urban environment. To overcome this issue, semi-supervised NMF (SEM-NMF) \cite{lee_semi-supervised_2010} has been proposed. It consists in decomposing, $\mathbf{W}_{F \times (K+J)}$ into two distinctive matrices: $\mathbf{W} = \left[ \mathbf{W_s}~\mathbf{W_r} \right]$ where $\mathbf{W_s}_{F \times K}$ is a fixed part of $\mathbf{W}$ composed of known audio spectra and $ \mathbf{W_r}_{F \times J}$, a mobile part which is updated, see eq. \ref{eq:W_r_SS}. Thus it is possible to let the method define the best elements to include in $\mathbf{W_r}$. Its dimension is set up as $J << K$ in order to consider, as a priority, the sound sources present in $\mathbf{W_s}$. $\mathbf{H}$ is then also decomposed in two matrices, $\mathbf{H}_{(K+J) \times N} = \genfrac[]{0pt}{0}{\mathbf{H_s}}{\mathbf{H_r}}$. Eq. \ref{eq:nmf} becomes

@@

$\mathbf{H_r}$ and $\mathbf{H_s}$ are updated separately, see eq. \ref{eq:H_r_SS} to eq. \ref{eq:H_s_SS}.

@@

In this study, $\mathbf{W_s}$ is composed of traffic audio spectra, as it is the sound source of interest. Sources included in $\mathbf{W_r}$ are other sound sources (corresponding to the interfering class) that can be present in the urban sound scenes. The traffic signal estimation is next defined by the fixed part,

@@

The addition of $\mathbf{W_r}$ gives more flexibility to the method to represent correctly the spectrogram $\mathbf{V}$. The representational capability is increased, thus the approach is more adaptive to the different urban sound environments. Applications of SEM-NMF can be found for musical \cite{weninger2012supervised, kitamura_music_2014} and speech contents \cite{joder2012real, mysore2011non}.

@@

With this operation, $\mathbf{W_0}$ is oriented to the focused sound source (the road traffic) but also can be adapted to the content of the scene thanks to the updates. After $N$ iterations, each element $k$ of the final dictionary, $\mathbf{W'}$, is compared with its initial value in $\mathbf{W_0}$, in order to identify which element is stayed close to the traffic component. A cosine similarity $D_{\theta}\left(\mathbf{W_0} \Vert \mathbf{W'} \right)$ is computed for each element $k$ as it is scale-invariant and bounded,

@@

where $\mathbf{w}$ is a $k$ element of $\mathbf{W}$ of $F \times 1$ dimensions. When $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$=1, the element $\mathbf{w'}$ is identical to $\mathbf{w_0}$. If $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$=1, both elements are fully different. The extraction of traffic elements in $\mathbf{W'}$ is carried out by a hard thresholding method \cite{donoho1994threshold}. It consists in weighting in a binary way $\mathbf{W'}$ according to $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$ and a threshold value $\alpha_k$ such as:

@@

To summarize, to approximate the audio spectrogram $\mathbf{V}$ and estimate the traffic component, 3 methods are used which deal differently with the \textit{a priori} knowledge on the traffic. SUP-NMF is only based on fixed \textit{traffic} elements and is constrained to used it to obtain $\mathbf{\tilde{V}}_{traffic}$. SEM-NMF completes this dictionary by the add of a mobile dictionary to better consider the interfering sound events and then offer more flexibility. Finally, TI-NMF considers first a dictionary composed of \textit{traffic} spectra, as SUP-NMF, but allows an update of them. The interest of an initiated dictionary is then to focus on the source of interest during the updates. This method therefore makes it more suitable for solving the generalization issues as it is the entire dictionary that can be fully adapted to the sound scenes. The elements that deviates too much from the originals spectra are then discarded by the thresholding step.

These above described methods are applied on an evaluation corpus, composed of simulated sound scenes, in order to compare the estimated traffic sound levels with an exact reference. This new sound corpus is designed by considering realistic urban sound environments with many kinds of mixed sound sources.

@@

\begin{figure*}[t]
\centering
\includegraphics[width=0.8\linewidth]{figures/realistic_urban_sound_scene_design.pdf}
\caption{Bloc diagram of the design of the evaluation corpus}
\label{fig:transcription}
\end{figure*}

The evaluation corpus must be both realistic and non ambiguous in terms of traffic sound level. The former calls for the use of real urban sound environments, and the latter imposes the use of controlled sound sources with known properties: onset and offset time location, and sound level. Indeed, the precise annotation of the sound level of a given source of interest in a real recording is a very complex problem. Two corpora are proposed to design a valid corpus of  evaluation. The first, called the reference corpus, is a set of recordings of urban sound environments, from which, what we call, a transcription is made, \textit{i.e.} which sound source is present from this time to this time. The second, called the elemental corpus, is a set of monophonic recordings of isolated sound sources that represent categories of events that are present in the reference scenes. The elemental corpus is built with still existing sounds and with specific recordings, see part \ref{part:simScene}. The evaluation corpus is then designed by replicating the reference corpus according to the transcription with the help of the elemental corpus, see Figure \ref{fig:transcription}.
More precisely, the reference corpus is composed of 76 recordings from 2 to 5 min, achieved in the \nth{13} district of Paris (France) at 19 different locations \footnote{Recordings were made as part of the Grafic project funded by Ademe}, see Figure \ref{fig:map_grafic}, which cover various sound environments. A complete description of the experimental protocol can be found in \cite{aumond_modelling_2017}. Two of the 76 recordings are rejected for the analysis because the audio files were corrupted, resulting in 74 valid audio files assumed as representative of the variety of urban sound environments. The recordings are listened and categorized within four different sound environments, as proposed in \cite{can_describing_2015}: \textit{park} (8 audio files with a cumulative duration of 16min01), \textit{quiet street} (35 audio files with a cumulative duration of 77min27), \textit{noisy street} (23 audio files with a cumulative duration of 56min10) and \textit{very noisy street} (8 audio files with a cumulative duration of 21min42). Then, each audio file is transcripted, noticing the start, end time and level of each sound event along with its sound class. This annotation phase makes it possible to produce simulated sound scenes with the same positions of sound events than the recordings and therefore as close as possible to the real scenes.

@@

\subsection{Generation of the evaluation corpus}\label{part:simScene}
The sound scenes are generated with the \textit{SimScene} software\footnote{Open-source project available at: \url{https://bitbucket.org/mlagrange/simscene}}, \cite{rossignol_simscene:_2015}, a simulation software generating monaural sound mixtures in wav format from an isolated sounds database.
This software has already been used in a wide range of experiments for sound detection algorithm assessment \cite{lafay_new_2014, benetos2016detection}. The \textit{SimScene} software allows the design of several type of sequencing, from \textit{abstract} ones (time indexes and amplitudes are drawn from random distributions) to  precise ones, where the time indexes and amplitudes for each event are set by the user. The latter type is considered in this study. As output, \textit{SimScene} generates an audio file of the global sound mixture as well as for each sound class present in the scene, see Figure \ref{fig:example_simScene}. It makes it possible to know their exact contributions in the scene, in particular their sound levels.

\begin{figure}[t]
    \centering
       \includegraphics[width=.85\linewidth]{./figures/exampleSimScene.pdf}
    \caption{Spectrogram of a simple scene created with the \textit{SimScene} software with one sound background (road traffic in blue) and 3 sound events (car horn in green, passing car in yellow and whistling bird in red).}
    \label{fig:example_simScene}
\end{figure}

To replicate the 74 recordings in simulated scenes, a high quality (wav format, 44.1 kHz sampling rate, high \textit{Signal Noise Ratio}) elemental corpus has been built-up from audio samples found online (\textit{freesound.org}) or with the help of an already existing sound database \cite{salamon2014dataset}. The elemental corpus is composed of two categories of sound: the \textit{event} category, which includes 245 brief sound samples considered as salient, with a 1 to 20 seconds duration and classified among 21 sound classes (\textit{ringing bell, whistling bird, car horn, passing car, hammer, barking dog, siren, footstep, metallic noise, voice \dots}) and the \textit{background} (or \textit{texture}) category gathering 154 long duration sounds ($\approx$ 1mn30), whose acoustic properties do not vary in time. This category includes among others \textit{whistling bird, crowd noise, rain, children playing in schoolyard, constant traffic noise} sound classes. Each sound class is composed of multiple samples (\textit{carHorn01.wav, carHorn02.wav \dots}) that are randomly chosen by the \textit{simScene} software to bring diversity.
As the road traffic is the main component in urban environment and is the sound source of interest, recordings of car passages have been made on the Ifsttar's runway. The recordings have been made for 4 different cars (Renault Scenic, Renault M\'egane, Renault Clio and Dacia Sandero), at different speeds and gear ratios. Overall, 103 car passages have been recorded. In order to avoid overfitting issues, the audio samples of the first two cars (Renault Scenic and Renault M\'egane) are included in the \textit{SimScene}'s elemental corpus (50 audio files in total). The last 53 audio samples are dedicated to the dictionary design, $\mathbf{W}$ as part of NMF, see section \ref{part:dictionary_building}. A full description of the recordings can be found in \cite{gloaguen_creation_2017}.

\begin{table}[t]
\centering
\caption{Mean \textit{Traffic Interfering Ratio} and its standard deviation for each sound environment.}
\begin{tabular}{lc}
 & $mTIR$ (dB)\\ \hline
 \textit{park} & -9.10 ($\pm$ 7.35) \\
 \textit{quiet street} & 0.88 ($\pm$ 5.92) \\
 \textit{noisy street} & 6.96 ($\pm$ 5.16) \\
 \textit{very noisy street} & 15.75 ($\pm$ 9.78) \\ \hline
\end{tabular}
\label{tab:mTIR}
\end{table}

With this built-up corpus, the \textit{SimScene} software and the transcriptions of the recordings, 74 simulated sound scenes are generated, which have the same temporal structure of the reference recordings. The sound level of each sound class is adjusted manually on each sound scene to be faithful compared to the recorded scenes. To check this adjustment, the mean \textit{Traffic Interfering Ratio} ($mTIR$) is calculated, see Table \ref{tab:mTIR}. It expresses, on all the scenes of a sound environment, the mean difference between the equivalent traffic sound levels of each scene, $L_{p,traffic}$, with the sound level of the \textit{interfering} sound class, $L_{p, interfering}$, which gathered all the other sound sources not related to the traffic, see Eq. \ref{eq:mTIR}. It quantifies the predominance of the traffic component for the 4 types of sound environments,
